---
title: "Data Rescue: Lin 2015"
format:
  html:
    toc: true
    eval: false #default to not evaluating the chunks, override this the template is completed
date: last-modified
date-format: YYYY-MMMM
bibliography:
  - Lin2015_Methods.bib
  - Lin2015.bib
authors:
  - id: ktoddbrown
    name:
      given: Katherine
      family: Todd-Brown
    orcid: 0000-0002-3109-8130
    affiliation:
      - ref: uf-ees
    role:
      - transcription: review
     # - standardization: lead #review
    #  - curation: lead #review

  - id: SavaScott
    name:
      given: Savannah
      family: Scott
    affiliation:
      - ref: uf-ees
    role:
      - transcription: lead
  #    - standardization: lead #review
   #   - curation: lead #review

  - id:  ChaseSaxour
    name:
      given: Chase 
      family: Saxour
    affiliation:
      - ref: uf-ees
    role:
      - transcription: secondary
    #  - standardization: lead #review
    #  - curation: lead #review
      

affiliations:
  - id: uf-ees
    name: University of Florida
    department: Environmental Engineering Sciences
    city: Gainesville
    state: Florida
    country: USA
    address: "365 Weil Hall \nP.O. Box 116580"
    postal-code: 32611
    url: https://essie.ufl.edu/ees/
    
timelog:
  - activity: 
      description: Primary Transcription and ReadMe Update
      who: SavaScott
      start: 2025-01-30
    #  end: 
      duration: P45M
      
  - activity: 
      description: Revise based on feedback
      who: ktoddbrown #github username
      #when you started and ended, can be just a year or down to the minute
      end: 2025-11-26
      #start: YYYY-MM-DD THH:MM
      #alternatively duration can be combined with a being/end
      duration: P30M
---


```{r}
#| label: setup
#| include: true
#| warning: false
#| message: false
#| echo: false


library(tidyverse) # data processing
library(kableExtra) # pretty tables
library(bibtex) # reading in BibTex files to R

# data files for level 0
methods.file <- "Lin2015_Methods.md"
table1.file <- 'Lin2015_Table1.csv'
table2.file <- 'Lin2015_Table2.cvs'

# Bibliograph files
primaryCitation.file <- 'Lin2015.bib'
methodsCitation.file <- 'Lin2015_Methods.bib'

# read function; you will need to create this function after you have your data
#... rescue reviewed and before you finish your data harmonization.
readsource.file <- '../../../R/readLin2015.R'
if(file.exists(readsource.file)){
  source(readsource.file)
}

#semantic files for level 1
of_variable.file <- '../../semantics/SoilDRaHVocabulary_of_variable.csv'
is_type.file <- '../../semantics/SoilDRaHVocabulary_is_type.csv'

```


# Data Summary

<!---- Fill in the full citation here, generally found on the journal page and link in the issue that you started on GitHub. Though this is at the top of the document, this should be one of the last things you do because a lot of this will be a direct copy-paste into the issue template. --->

Lin, Y., Scarlett, R.D. & King, J.Y. Effects of UV photodegradation on subsequent microbial decomposition of Bromus diandrus litter. Plant Soil 395, 263–271 (2015). https://doi.org/10.1007/s11104-015-2551-0

For the discussion of this data rescue see the Github issue: https://github.com/ktoddbrown/SoilDRaH/issues/185.

<!--- After you read through the paper write a brief summary of what the study
measured and what the data was used for in. This should be <250 words.--->

Lin2015 primarily measures ... 
They use this data to look at ...

## Tables and figures

<!--- Write a brief discription of all the tables and figures in this source. This should include full variable names when possible. It is perfectly fine to restate the captions here or copy other text that feels complete. Please keep this to a single sentence for each table/figure.--->

- Table 
  1) Table 1. Bromus diandrus litter samples and their field UV exposure characteristics
  2) Table 2. Effects of UV treatment on chemical characteristics of the litter

- Figure 
  1) Effects of UV manipulation and exposure periods on microbial respiration in the litter.
  2) Figure 2. Subsequent microbial respiration from the litter as a function of time for litter exposed during summer.
  3) Figure 3. Relationship between litter N concentration and cumulative microbial respiration during the periods.
  4) Figure 4. Conceptual model of solar radiation effect on little mass loss


## Fit for purpose: ProjectName

<!--- Write a brief discription of the project you are processing this data source for. This could be a copy-paste from other documentation. ---->

This data is identified as a data source for PROJECT.
The purpose of this PROJECT is...

<!--- Include a list of the key variables identified for the project. This is used to identify the elements for the paper for the data rescue. An example is below for a basic geo-located soil carbon stock project. --->

  - Location: The geo-location is given in the site description section of the methods. Both latitude and longitude are given but not the datum/projection that was used.
  - Soil carbon stock: Soil carbon stock was calculated from organic carbon fraction and bulk density and is reported in table 3 across two treatments and one control. Coarse fraction was not reported and does not appear to be included.
  
## Files

<!--- Include a brief explaination to all the files in the folder--->

These files are in the Lin2015 data rescue.

- [Readme](README_Lin2015.qmd)
  + This is the primary file that documents the transcriptions and decision made during data rescue.
- [Lin2015.bib](Lin2015.bib)
  + citation for article transcribed
- [Lin2015_Methods.bib](Lin2015_Methods.bib)
  + citations for the methods section of the article
- [Lin2015_Methods.md](Lin2015_Methods.md)
  + methods transcribed from primary article
- [Lin2015_TableN.csv](Lin2015_TableN.csv)
  + table N from primary article with ...

# Data Rescue working notes 

## Data Rescue Plan

+ Lead
  - [x] Read the article or data resource and the discription of the target project
  - [x] Create a fresh fork of the primary repository https://github.com/ktoddbrown/SoilDRaH/tree/main, it might be useful to rename it 'SoilDRaH_Lin2015'
  - [x] Create a copy of this template folder and rename it to Lin2015
  - [x] Fill out Data summary section (except for the Files), the Data Rescue Plan (here!), and update the author section of the yml to reflect who is working on the rescue
  - [x] Create a GitHub issue from the Data Rescue template and get feedback from a reviewer on the project (Dr Todd-Brown or a grad student typically).
  - [x] Add the reviewer to the author list in the yml
+ Do
  - [ ] Create a spreadsheet with TableN and export it to csv
  - [ ] Extract points from FigureM and export it to csv
  - [x] Transcribe or copy the method section
  - [x] Insert references to bib entries
  - [x] reformat equations and mathematical symbols as [LaTex](https://en.wikibooks.org/wiki/LaTeX/Mathematics)
  - [x] reformat text using [Markdown](https://www.markdownguide.org/cheat-sheet/)
  - [x] Create a bib file with the citations from the methods section
  - [ ] Push to your local fork and get a second transcription pushed to your local fork by someone else on the project
+ Measure
  - [ ] cross check with second transcriber for tables and methods, and resolve any differences
  - [ ] compile ReadMe with all chunks up through the Level 0 data rescue section with `eval: true`
  - [ ] push to your local fork and create a pull request to the main branch of the primary repository
  - [ ] address any revisions suggested in your pull request review
  - [ ] celebrate!

## Table N

<!--- Common table modifications include: inserting an extra column to pivot sub-headers, filling in NA to identify blanks, and adding Latex or Markdown to denote formatting.---->

Table N was modified ....

```{r}
#| label: table1Check
#| eval: false #remove when you have the file transcribed


read_csv(file = tableN.file, #read in the specific file
         skip = 1, #skip the first line with the caption
         #read everything as a character, don't worry about formatting yet
         col_types = cols(.default = col_character())) |>
  kable(caption = read_csv(file = tableN.file, #read in that file again but...
                           n_max = 1, #only read the first line
                           col_names = 'caption', #call it the caption
                           #read everything as a character
                           col_types = cols(.default = col_character()),
                           show_col_types = FALSE #do not return messages
                           )$caption #and take it out of the table
  )
```

## Methods

```{bash}
#| label: compareTranscriptions
#| eval: false
# You will want to repeat the line below to cross check each csv and md file in the
#... data rescue. Having no lines flagged by this command will ensure that each character matches exactly.
diff --strip-trailing-cr --suppress-common-lines -y Lin2015_Methods.md Lin2015_Methods_CS.md
```

{{< include Lin2015_Methods.md >}}

<!----
## Citation notes from Methods

Below are the citation notes from the methods section.
This generally doesn't read well in the final report so commenting it out here but it is very useful for tracking citations when you are doing the transcriptions.
Some citations are missing, those are noted here and how best guess were generated.
Think of this as a lab notebook documenting how you found each citation.

- [ ] short citation
  + copy from bib
  + manually entered or direct export
  
- [x] Lin Y, King J (2014) Effects of UV exposure and litter position on decomposition in a California grassland. Ecosystems 17:158–168
  + copy from bib

- [x] Van Soest PJ (1963) Use of detergents in the analyses of fibrous feeds. A rapid method for the determination of fiber and lignin. J Assoc Off Anal Chem 46:829–835
  + copy from bib
  
- [x] Austin AT, Vivanco L (2006) Plant litter decomposition in a semi-arid ecosystem controlled by photodegradation. Nature 442:555–558
  + copy from bib
  
- [x] Brandt LA, King JY, Hobbie SE, Milchunas DG, Sinsabaugh RL (2010) The role of photodegradation in surface litter decomposition across a grassland ecosystem precipitation gradient. Ecosystems 13:765–781
  + copy from bib
  
- [x] Rozema J, Tosserams M, Nelissen HJM, van Heerwaarden L, Broekman RA, Flierman N (1997) Stratospheric ozone reduction and ecosystem processes: enhanced UV-B radiation affects chemical quality and decomposition of leaves of the dune grassland species Calamagrostis. Plant Ecol 128:285–294
  + copy from bib

- [x] von Lützow M, Kögel-Knabner I, Ekschmitt K, Flessa H, Guggenberger G, Matzner E, Marschner B (2007) SOM fractionation methods: relevance to functional pools and to stabilization mechanisms. Soil Biol Biochem 39:2183–2207
  + copy from bib

---->

# Level 0 Data rescue

This level 0 data read parses the transcribed files with minimal cleaning or transformations.

```{r}
#| label: readLevel0
#This chunk has two purposes. 
#...1) Check the formatting by reading in everything 
#...2) Create a list of everything to process later in level 1

data.lvl0.ls <- list(
  #Read in a list of all the bib files
  citation = list(
    #Citation for the article transcriptions are pulled from
    primary = read.bib(file = primaryCitation.file), 
    #Citations for all referenced articles
    methods = read.bib(file = methodsCitation.file)
  ),
  #Read in the text transcription of the article's methods section
  method = read_lines(file = methods.file),
  #Read in the results as tables or figure transcriptions. This includes
  #...the caption as well as the tables themselves
  data = list(
    Table1 = list(
      #Read the caption as a text string. Captions are the first cell on 
      #...the first row.
      caption = read_csv(file = tableN.file,
                         col_types = cols(.default = col_character()),
                         n_max = 1, col_names = FALSE)$X1[1],
      #Read in all the data, skipping the first row with the caption and read
      #...in the table as character. This element is a tibble (data.frame).
      primary = read_csv(file = tableN.file,
                         col_types = cols(.default = col_character()),
                         skip = 1)
    ) #End 'Table1' element
    #Repeat for all tables and figures...
   ) #end 'data' element
)

```

The code is also in `readsource.file` and returned with the `dataLevel = 'level0'` argument.
Below verifies that the data from the function matches what is coded above.

```{r}
#| label: checkReadFunction_lvl0
if(file.exists(readsource.file)){
  datafunction.lvl0.ls <- readLin2015(dataDir = '.',
                                         dataLevel = 'level0')
  
  if(isTRUE(all.equal(datafunction.lvl0.ls, data.lvl0.ls))){
    print('Read function matches code here.')
  }else{
    print('There is a mismatch in the data objects between the code here and the read function.')
  }
}else{
  print('The read function does not exist.')
}
```


# References

<!---- This space will be filled with the citations when rendered. Leave this blank. ---->

